Elastic Scattering in General Relativistic Ray Tracing for Neutrinos

### 2018-06-15 paper.tex
Update oscillation potentials plots s->r and cm->km

The formalism of the single-angle approximation requires knowledge of
the unoscillated neutrino contribution to the Hamiltonian matrix
along a given test trajectory.
This is a function of the unoscillated neutrino self-interaction potential,
which for the $\alpha$-th flavor is:
\begin{equation}
  \label{eqn:V_nu_alpha}
  V_{\nu_\alpha,0}(\varepsilon,\ell_\mu) =
  \frac{\sqrt{2}G_{\rm F}}{(2\pi)^3} \varepsilon^2
  \oint d\Omega'(1-\varpi')f^{\nu_\alpha}(\varepsilon,\ell'_\gamma),
\end{equation}
with the test ray propagating in direction $\ell_\mu$,
ambient rays propagating in directions $\ell'_\gamma$,
and the cosine of the angle between these given by
$\varpi'=\psi^{\mu\gamma} \ell_\mu \ell'_\gamma$.
As in the moment equations (Eqns.~\ref{eqn:J}-\ref{eqn:Ka}),
the integral is taken over all directions $\ell'_\gamma$.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{20171223-potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg-noscat}
  % Plot:   $DIR/RunBA/RunBA-025deg-noscat-production01/potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg.pdf
  % Data:   $DIR/000Mo-025deg-*km/Lev2-e30a200b30-noscat-04
  % Script: $DIR/nu_potentials-production.nb
  % Config: $DIR/RunBA/RunBA-025deg-noscat-production01/nu_potentials-config-production.nb
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/
  \caption{Neutrino oscillation potentials
    for the \emph{noscat} case
    along a test trajectory
    originating at the surface of the hypermassive neutron star, and
    proceeding outward along a radial coordinate ray with angle
    $\theta=25^{\circ}$ with respect to the polar axis.
    The trajectory is parameterized by the coordinate length $s$.
    We plot the vacuum potential due to mass-squared differences (gray bands),
    the matter potential $V_e$ due to forward scattering on $e^-$ and $e^+$
    (dark green),
    and the self-interaction potential $V_{\nu\nu,0}$ due to forward
    scattering on ambient neutrinos (light green, solid where positive,
    dashed where negative);
    additionally we plot the $\nu_e$ and $\bar{\nu}_e$ components composing
    the self-interaction potential (light blue and dark blue respectively).
    The two gray bands at $V\sim10^{-22}$~erg and $V\sim10^{-24}$~erg
    indicating the two vacuum energy scales for 10--30~MeV neutrinos,
    set the positions of possible Mikheyev-Smirnov-Wolfenstein (MSW)
    or nutation resonances \cite{malk2012-mnr_1}.
    Inside the symmetric point at $s\sim100\,{\rm km}$ the total unoscillated
    self-interaction potential $V_{\nu\nu,0}$ is positive,
    i.e.\ $\nu_e$-dominated.
    }
  \label{fig:V_nunu-noscat}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{20171223-potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg-scat}
  % Plot:   $DIR/RunBA/RunBA-025deg-scat-production01/potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg.pdf
  % Data:   $DIR/000Mo-025deg-*km/Lev2-e30a200b30-scat-04
  % Script: $DIR/nu_potentials-production.nb
  % Config: $DIR/RunBA/RunBA-025deg-scat-production01/nu_potentials-config-production.nb
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/
  \caption{Same as Fig.~\ref{fig:V_nunu-noscat},
    but for the \emph{scat} case.
    Unlike the case with scattering turned off,
    here the total unoscillated self-interaction potential $V_{\nu\nu,0}$
    is everywhere negative, i.e.\ $\bar{\nu}_e$-dominated.
    }
  \label{fig:V_nunu-scat}
\end{figure}

@@

In Fig.~\ref{fig:survival-noscat}, we report the survival probabilities
from the calculation with elastic scattering turned off.
Using neutrino mixing angle $\theta_{12}$ and the inverted hierarchy
(the normal hierarchy gives qualitatively similar results)
electron neutrinos and antineutrinos start to oscillate around
700~km in the form of a collective neutrino oscillation,
causing both $\nu_e$ and $\bar{\nu}_e$ to convert to heavy lepton neutrinos
and antineutrinos respectively.
A similar effect was seen in \cite{fren2017-flavor_bns,tian2017-flavor_bns}.
In Fig.~\ref{fig:survival-scat}, we report the survival probabilities
of an otherwise identical calculation, but with elastic scattering turned on.
In this case, electron neutrinos and antineutrinos start to oscillate around
400~km in the form of a standard matter neutrino resonance:
at first both $\nu_e$ and $\bar{\nu}_e$ convert to heavy lepton neutrinos,
but as the transformation progresses, the $\bar{\nu}_e$ partially
return to their original flavor.

>>>>

The formalism of the single-angle approximation requires knowledge of
the unoscillated neutrino contribution to the Hamiltonian matrix
along a given test trajectory.
This is a function of the unoscillated neutrino self-interaction potential,
which for the $\alpha$-th flavor is:
\begin{equation}
  \label{eqn:V_nu_alpha}
  V_{\nu_\alpha,0}(\varepsilon,\ell_\mu) =
  \frac{\sqrt{2}G_{\rm F}}{(2\pi)^3} \varepsilon^2
  \oint d\Omega'(1-\omega')f^{\nu_\alpha}(\varepsilon,\ell'_\gamma),
\end{equation}
with the test ray propagating in direction $\ell_\mu$,
ambient rays propagating in directions $\ell'_\gamma$,
and the cosine of the angle between these given by
$\omega'=\psi^{\mu\gamma} \ell_\mu \ell'_\gamma$.
As in the moment equations (Eqns.~\ref{eqn:J}-\ref{eqn:Ka}),
the integral is taken over all directions $\ell'_\gamma$.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{20180615-potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg-noscat}
  % Plot:   $DIR/RunBA/RunBA-025deg-noscat-production01/potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg.pdf
  % Data:   $DIR/000Mo-025deg-*km/Lev2-e30a200b30-noscat-04
  % Script: $DIR/nu_potentials-production.nb
  % Config: $DIR/RunBA/RunBA-025deg-noscat-production01/nu_potentials-config-production.nb
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/
  \caption{Neutrino oscillation potentials
    for the \emph{noscat} case
    along a test trajectory
    originating at the surface of the hypermassive neutron star, and
    proceeding outward along a radial coordinate ray with angle
    $\theta=25^{\circ}$ with respect to the polar axis.
    The trajectory is parameterized by the coordinate radius $r$.
    We plot the vacuum potential due to mass-squared differences (gray bands),
    the matter potential $V_e$ due to forward scattering on $e^-$ and $e^+$
    (dark green),
    and the self-interaction potential $V_{\nu\nu,0}$ due to forward
    scattering on ambient neutrinos (light green, solid where positive,
    dashed where negative);
    additionally we plot the $\nu_e$ and $\bar{\nu}_e$ components composing
    the self-interaction potential (light blue and dark blue respectively).
    The two gray bands at $V\sim10^{-22}$~erg and $V\sim10^{-24}$~erg
    indicating the two vacuum energy scales for 10--30~MeV neutrinos,
    set the positions of possible Mikheyev-Smirnov-Wolfenstein (MSW)
    or nutation resonances \cite{malk2012-mnr_1}.
    Inside the symmetric point at $r\sim100\,{\rm km}$ the total unoscillated
    self-interaction potential $V_{\nu\nu,0}$ is positive,
    i.e.\ $\nu_e$-dominated.
    }
  \label{fig:V_nunu-noscat}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{20180615-potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg-scat}
  % Plot:   $DIR/RunBA/RunBA-025deg-scat-production01/potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg.pdf
  % Data:   $DIR/000Mo-025deg-*km/Lev2-e30a200b30-scat-04
  % Script: $DIR/nu_potentials-production.nb
  % Config: $DIR/RunBA/RunBA-025deg-scat-production01/nu_potentials-config-production.nb
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/
  \caption{Same as Fig.~\ref{fig:V_nunu-noscat},
    but for the \emph{scat} case.
    Unlike the case with scattering turned off,
    here the total unoscillated self-interaction potential $V_{\nu\nu,0}$
    is everywhere negative, i.e.\ $\bar{\nu}_e$-dominated.
    }
  \label{fig:V_nunu-scat}
\end{figure}

@@

Fig.~\ref{fig:survival-noscat} shows the survival probabilities
from the calculation with elastic scattering turned off.
Using neutrino mixing angle $\theta_{12}$ and the inverted hierarchy
(the normal hierarchy gives qualitatively similar results)
electron neutrinos and antineutrinos start to oscillate around
700~km in the form of a collective neutrino oscillation,
causing both $\nu_e$ and $\bar{\nu}_e$ to convert to heavy lepton neutrinos
and antineutrinos respectively.
A similar effect was seen in \cite{fren2017-flavor_bns,tian2017-flavor_bns}.
Fig.~\ref{fig:survival-scat} shows the survival probabilities
of an otherwise identical calculation, but with elastic scattering turned on.
In this case, electron neutrinos and antineutrinos start to oscillate around
400~km in the form of a standard matter neutrino resonance:
at first both $\nu_e$ and $\bar{\nu}_e$ convert to heavy lepton neutrinos,
but as the transformation progresses, the $\bar{\nu}_e$ partially
return to their original flavor.
